\textbf{How open are hybrid journals included in transformative agreements?}

### 2024-02-12 analysis/manuscript.Rmd
refs

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
csl: apa.csl
bibliography: references.bib
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300,
  dev = c("png", "pdf"),
  fig.path = "fig/"
)
options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

@@

@@

This study combines data from multiple publicly available data sources as diagrammatically shown in Figure \ref{fig:data_workflow}. 
Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool provided information about journal portfolios and participating institutions. 
After identification of hybrid journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to determine open access availability on publisher websites.
To identify open access articles published through transformative agreements, first author affiliations from OpenAlex were subsequently linked to eligible institutions according to the transformative agreement data.

```{r data_workflow, echo=FALSE, fig.cap="Data collection workflow to obtain article-level transformative agreement data"}
knitr::include_graphics(here::here("analysis", "data_collection_workflow.pdf"))
```

@@

Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool.[^checkertool] 
The dump consisted of multiple online Google spreadsheets where each data file represented one agreement listed in the ESAC Transformative Agreement Registry.[^esacregistry]
From the retrieved spreadsheet files, journals and institutions involved per agreement were obtained.

A limitation of using the Journal Checker Tool to study the development of transformative agreements over time is that it only provided data about active transformative agreements.
As a result, expired transformative agreements were constantly removed from the publicly available data dump. 
To address this, four different snapshots were safeguarded and combined for this study: self-archived versions from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
This ensured that transformative agreements, which ended from 2021 onwards, were included.
<!-- Agreements ending before 2021 could not be obtained.-->
Overall, the four combined Transformative Agreement Data dumps used in this study contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` out of `r length(unique(esac_df$ID))` agreements listed in the ESAC registry by December 2023.

The Transformative Agreement Data dumps link agreements to journals represented by journal names and ISSN. 
After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by the ISSN International Centre, journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
Because transformative agreements can include both full open access and hybrid journals, the data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals [@goldoa]. 
As shown in Figure \ref{fig:method_fig}A, combining different data sources considerably extended the journal matching to exclude full open access journals from transformative agreements.
In total, `r n_oa_journals` full open access journals were excluded based on ISSN matching.
The overlap between the three data sources was `r round(length(unique(Reduce(intersect, my_venn))) / length(unique(Reduce(union, my_venn)))* 100,0)`%.
The Gold OA journals dataset alone added `r length(unique(Reduce(setdiff, my_venn)))` journals, while the DOAJ comprised `r length(setdiff(my_venn$DOAJ, union(my_venn[["GOLD OA"]], my_venn$OpenAlex)))` full open access journals not listed in either of the other two sources. 
These full open access journals were mostly launched in 2022.

@@

After identifying hybrid journals included in transformative agreements, article metadata was retrieved from Crossref<!. <!-- and mapped to OpenAlex for obtaining author metadata including affiliations. --> November 2023 database snapshot for the five-year period 2018 to 2022. 
Because Crossref metadata lacked information to distinguish between original research articles including review and other types of journal content, which are often not covered by transformative agreements [@Borrego_2020], only articles published in regular issues indicated by non-numeric pagination were included.
Furthermore, an expanded version of Unpaywall’s paratext recognition approach was applied to exclude non-scholarly journal content such as table of contents.

Open access articles in hybrid journals were identified through Creative Commons (CC) license information in Crossref metadata. 
License information relative to the "accepted manuscript (AM)" version were not considered. 
Crossref was used for open access identification because transformative agreements workflows generally require publishers to deliver CC license information to this DOI registration agency [@Geschuhn_2017].
Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, not all publisher provide such license information a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], demonstrates the ongoing challenges to identify open access articles in hybrid journals [@Mart_n_Mart_n_2018; @Jahn_2021].
Overall, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" in OpenAlex. 
The biggest differences concerned articles published between 2018 and 2020.
In 2022, however, Crossref and OpenAlex open access numbers only differ slightly (`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` records using Crossref vs. `r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` in OpenAlex).
Upon inspection, notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society and Emerald.
Crossref license metadata was more complete with regard to articles from the publisher Wiley and American Chemical Society.
Finally, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023]. According to the OpenAlex release notes, fixing this issue was still ongoing, which might also explain this discrepancy.

@@

Author metadata are crucial for the planning and evaluation of transformative agreements [@Schimmer_2015; @Geschuhn_2017], in particular information about corresponding authors who are generally responsible to arrange open access publication [@Borrego_2020]. 
Here, affiliations were retrieved from OpenAlex.
However, because of low coverage, this study focused on first authors and their affiliations.
First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016]. 
Overall, around 90% of studied articles had first author affiliation metadata in OpenAlex, whereas the coverage of articles with corresponding author information was around 54%.

@@

To assess the impact of transformative agreements to hybrid open access, participating institutions from the Transformative Agreement Data dump, which were crowd-sourced from the agreements and consortia that successfully negotiated an agreement, were matched with first author affiliations recorded by OpenAlex using the ROR-ID.
The matching also took into account the duration of an agreement according to the ESAC registry.
Upon inspection, Transformative Agreement Data did not cover associated institutions comprehensively like university hospitals or institutes of large research organisations like the Max Planck Society. To improve the matching, Transformative Agreement Data was automatically enriched with associated institutions using OpenAlex's insitution data.

In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022 (see Figure \ref{fig:method_fig}. 
Hybrid journals included in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of total global output over the same time period according to Crossref, while full open access journals accounted for
`r cr_stats |> filter(src == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`% of total article volume.

@@

Throughout this mostly automated data gathering and analysis process, we used tools from the Tidyverse (Wickham et al., 2019) for the R programming language (R Core Team, 2020). 
The resulting data is openly available through an R package, hoaddata. 
Following @Marwick_2018, hoaddata contains not only the datasets used in the data analysis. It also includes codeused to compile the data by connecting it the cloud-based Google Big Query data warehouse, where the big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
To increase the computational reproducibility, the R package was automatically built using GitHub Actions, a continuous integration service.

@@

```{r}
#| fig-cap: "Relative growth of open access in hybrid journals through transformative agreements between 2018 and 2022 per publication year  according to the total number of articles indexed in Crossref. The blue areas represent open access through transformative agreements, as indicated by the affiliation with institutions that had a corresponding transformative agreement in place at the time of publication according to the first author affiliation data in OpenAlex matched with journal and institutional data from cOAlition S Transformative Agreements Public Data. Grey areas depict open access articles where no link to an agreement could be established. (A) Proportion of open access articles in hybrid journals per year. (B) Percentage of hybrid open access via agreements per year. Boxplots show the proportion of open access articles by individual hybrid journals (C) and individual open access uptake rates by individual hybrid journals and open access funding (D) per publication year. Horizontal lines correspond to the lower quartile, median, and upper quartile, with whiskers extending to ±1.5 of interquartile range. The individual outliers are not shown. Note that data on transformative agreements ending before June 2021 were not available at the time of this study."

@@

Figure 2A shows a moderate growth in the proportion of open access articles in hybrid journals, comparing the overall open access uptake and the impact of transformative agreements on this trend.
Over the five-years period from 2018 to 2022, open access increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`). 
At the same time, the total article volume of the investigated journals grew from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.

Figure 2B highlights that the majority of open access articles in hybrid journals were made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
However, there was also a notable growth in open access provision through individual publication fees, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`). This suggests that publishers were able to gain equally from individual and institutional open access publishing options.

Figure 2C depicts the substantial variations among the hybrid journals included in transformative agreements in terms of open access uptake.
Although the median generally follows the trend shown in Figure 2A, the farther stretch of upper quartiles and whiskers over the years illustrates that an increasing number of journals published an above-average proportion of open access articles. 
In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a Creative Commons license in the same year. 
These journals were, on average, smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`). 
Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access) and Bioinformatics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "1367-4803") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "1367-4803") |> pull(n)` articles in 2022 were open access), which flipped to full open access as of January 2023.

When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements substantially contributed to the growth of open access over the years (Figure 2D). 
Examples of such journals include those with a scope on specific countries or regions, where also transformative agreements were implemented. 
For instance, in 2022, the Germany-based journals Zeitschrift für Erziehungswissenschaft and Zeitschrift für Politikwissenschaft, as well as the Scandinavian Political Studies adressing the Nordic countries, achieved an overall open access uptake of more than 90% just through transformative agreements. 
Despite the rise in transformative agreements, it is worth noting that other means of publishing open access in hybrid journals remained common. 
In total, `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_no_ta") |> nrow()` journals published open access articles from authors affiliated with institutions without transformative agreements in place, while `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_ta_oa") |> nrow()` journals published at least one open access article through a transformative agreement in the same year.

@@

```{r}
publisher_league_tbl |>
  mutate(across(contains(c("perc", "cum")), ~ paste(round(., 3) * 100))) |>
  kbl(
    booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE, position = "H",
    caption ="Hybrid open access through transformative agreements market shares 2018-2022",
    col.names = c(
      "Publisher",
      "Total",
      "%",
      "C%",
      "Total",
      "%",
      "C%",
      "Total",
      "%",
      "C%",
      "Total",
      "%",
      "C%"
    )
  ) |>
  kable_styling(font_size = 10, position = "center") |>
  add_header_above(c(
    " " = 1,
    "Hybrid journals" = 3,
    "Articles" = 3,
    "OA articles" = 3,
    "TA OA articles" = 3
  ))
```

>>>>

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
    extra_dependencies: ["flafter"]
csl: apa.csl
bibliography: references.bib
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300,
  dev = c("png", "pdf"),
  fig.path = "fig/",
  fig.pos = "ht!", out.extra = ""
)
options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

@@

For over two decades, hybrid open access journal publishing, which makes some articles freely available under an open license while others remain behind a paywall, has been discussed as a  model for transitioning the subscription system to full open access [@Prosser_2003].
The idea was that when journals publish more and more open access articles, they could reduce revenues from subscriptions, while libraries and funders could change their funding models and shift expenditures from subscription to open access. 
However, initial approaches, mainly based on publication fees, also called article processing charges (APCs), did not contribute substantially to an large open access uptake. 
In 2012 ddd - at the same time

Tthe introduction of transformative agreements, a collective term for agreements starting from mid-2015, has resulted in a new push for hybrid open access. In particular, the Max Planck Society's Berlin 2020 initiative and plan has advocated these types of agreements as a transitional means to progress with open access. As a result, more than 800 transformative agreements have been tracked so far, which have been instrumental in funding activities and controversies. today, all publsiehrs offer such agreements

Advocates of these agreements argue that they not only overcome the issue of double-dipping but also help to shift library spending from subscriptions to publications, streamline expenditures, and offer more value than big deals, which focus on large publishers.  A few agreements can impact an institution's open access share, resulting in a significant proportion of open access content. increase transparency

However, opponents of transformative agreements perceive continued reliance on big deals as problematic. They argue that the focus on large publishers perpetuates the business relationships that led to the serial crisis and lacks evidence that this leads to a reduction of In addition, concerns over inequality abound. could not benn mass removal

Effects can be seen, mixed consequences. some countries who have introduced this fundign effects netherlands, sweden. no in germany, elsevier deal to price reduction and first steps towards mroe privacy. however, main goals associated woth trnasofrtiomative, the transition of the journal landscape not achivned, publsiehrs were not able to meet growtht raget, sweden did not see the flipping and has

This study contributes

@@

This study combines data from multiple publicly available data sources as diagrammatically shown in Figure \ref{fig:data_workflow}. 
Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool provided information about journal portfolios and participating institutions. 
After identification of hybrid journals by excluding full open access journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
To determine open access articles published through transformative agreements, first author affiliations from OpenAlex [@priem2022openalex] were subsequently linked to eligible institutions according to the transformative agreement data.

```{r data_workflow, echo=FALSE, fig.cap="Data collection workflow"}
knitr::include_graphics(here::here("analysis", "data_collection_workflow.pdf"))
```

@@

Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool.[^checkertool] 
The dump consists of multiple online Google spreadsheets where each data file represents one agreement listed in the ESAC Transformative Agreement Registry.[^esacregistry]
From the retrieved spreadsheet files, journals and institutions involved per agreement were obtained.

A limitation of using the Journal Checker Tool and its underlying publicly available data dump to study the development of transformative agreements over time is that expired transformative agreements are constantly removed. 
To address this, four different snapshots were safeguarded and combined for this study: self-archived versions from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
This ensured that transformative agreements, which ended from 2021 onwards, were included, representing the majority of transformative agreements.
<!-- Agreements ending before 2021 could not be obtained.-->
Overall, the four combined Transformative Agreement Data dumps used in this study contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` out of `r length(unique(esac_df$ID))` agreements listed in the ESAC registry by December 2023.

The Transformative Agreement Data dumps link agreements to journals represented by journal names and ISSN. 
After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by the ISSN International Centre, journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry. Furthermore, journal subjects according to the All Science Journal Classification code (ASJC) were added from the Scopus journal source list as of August 2023.

Because transformative agreements can include both full open access and hybrid journals, the data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals [@goldoa]. 
As shown in Figure \ref{fig:method_fig}A, combining different data sources considerably extended the journal matching to exclude full open access journals from transformative agreements.
In total, `r n_oa_journals` full open access journals were excluded based on ISSN matching.
The overlap between the three data sources was `r round(length(unique(Reduce(intersect, my_venn))) / length(unique(Reduce(union, my_venn)))* 100,0)`%.
The Gold OA journals dataset alone added `r length(unique(Reduce(setdiff, my_venn)))` journals, while the DOAJ comprised `r length(setdiff(my_venn$DOAJ, union(my_venn[["GOLD OA"]], my_venn$OpenAlex)))` full open access journals not listed in either of the other two sources. 
These full open access journals were mostly launched in 2022.

@@

After identifying hybrid journals included in transformative agreements, article metadata was retrieved from the Crossref <!-- and mapped to OpenAlex for obtaining author metadata including affiliations. --> November 2023 database snapshot for the five-year period 2018 to 2022 according to the issued date, representing the earliest known publication date. 
Because Crossref metadata lacked information to distinguish between original research articles including review and other types of journal content, which are often not covered by transformative agreements [@Borrego_2020], only articles published in regular issues indicated by non-numeric pagination were included.
Furthermore, an expanded version of Unpaywall’s paratext recognition approach was applied to exclude non-scholarly journal content such as table of contents.

Open access articles in hybrid journals were identified through Creative Commons (CC) license information in Crossref metadata. 
License information relative to the "accepted manuscript (AM)" version were not considered. 
Crossref was used for open access identification because transformative agreements workflows generally require publishers to deliver CC license information to this DOI registration agency [@Geschuhn_2017].
Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], highlighted ongoing challenges to identify hybrid open access [@Mart_n_Mart_n_2018; @Jahn_2021; @Zhang_2022; @Butler_2023].
Here, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" according to the OpenAlex November 2023 release, which was used throughout this study.
The biggest differences concerned articles published between 2018 and 2020.
In 2022, however, Crossref and OpenAlex open access numbers only differ slightly (`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` records using Crossref vs. `r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` in OpenAlex).
Notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society and Emerald.
Crossref license metadata was more complete with regard to articles from the publisher Wiley and American Chemical Society.
Finally, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023]. According to the OpenAlex release notes, fixing this issue was still ongoing, which might also explain this discrepancy.

@@

Affiliation metadata about corresponding authors are crucial for the planning and evaluation of transformative agreements, because they are considered to be responsible to arrange open access publication [@Schimmer_2015; @Geschuhn_2017; @Borrego_2020]. 
Here, country and institutional affiliations were retrieved from OpenAlex.
However, because of low coverage in OpenAlex, this study focused on first authors and their affiliations instead.
First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016], and can be therefore assumed as a proxy to measure to open access payments [@Zhang_2022].
Overall, around 90% of studied articles had first author affiliation metadata in OpenAlex, whereas the coverage of articles with corresponding author information was around 54%.

@@

To assess the impact of transformative agreements to hybrid open access,
participating institutions from the Transformative Agreement Data dump,
which were crowd-sourced from the agreements and consortia that
successfully negotiated an agreement, were matched with first author
affiliations recorded by OpenAlex using the ROR-ID.
The matching also took into account the duration of an agreement according to the ESAC registry.
Upon inspection, Transformative Agreement Data did not cover associated institutions comprehensively like university hospitals or institutes of large research organisations like the Max Planck Society. To improve the matching, Transformative Agreement Data was automatically enriched with associated organisations using OpenAlex's institution entity.

In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022 (see Figure \ref{fig:method_fig}B). 
Hybrid journals included in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of total global output over the same time period according to Crossref, while full open access journals accounted for
`r cr_stats |> filter(is.na(src) & oa_status == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`% of total article volume.

@@

Throughout this mostly automated data gathering and analysis process, we used tools from the Tidyverse (Wickham et al., 2019) for the R programming language (R Core Team, 2020). 
The resulting data is openly available through an R package, hoaddata. 
Following @Marwick_2018, hoaddata contains not only the datasets used in the data analysis. It also includes code used to compile the data by connecting it the cloud-based Google Big Query data warehouse, where the big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
To increase the computational reproducibility, the R package was automatically built using GitHub Actions, a continuous integration service.

@@

```{r results_overview}
#| fig-cap: "Relative growth of open access in hybrid journals in transformative agreements between 2018 and 2022 per publication year. The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established (according to matching OpenAlex first author affiliations matched with cOAlition S transformative agreement data). (A) Proportion of open access articles in hybrid journals per year. (B) Percentage of hybrid open access via agreements per year. Boxplots show the proportion of open access articles by individual hybrid journals (C) and individual open access uptake rates by individual hybrid journals and open access funding (D) per publication year. The individual outliers are not shown. Note that data on transformative agreements ending before June 2021 were not available for this study."

@@

Figure \ref{fig:results_overview}A shows a moderate growth in the proportion of open access articles in hybrid journals, comparing the overall open access uptake and the impact of transformative agreements on this trend.
Over the five-years period from 2018 to 2022, open access increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`). 
At the same time, the total article volume of the investigated journals grew from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.

Figure \ref{fig:results_overview}B highlights that the majority of open access articles in hybrid journals were made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
However, there was also a notable growth in open access provision through individual publication fees, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`). This suggests that publishers were able to gain equally from individual and institutional open access publishing options.

Figure \ref{fig:results_overview}C depicts the substantial variations among the hybrid journals included in transformative agreements in terms of open access uptake.
Although the median generally follows the trend shown in Figure \ref{fig:results_overview}A, the farther stretch of upper quartiles and whiskers over the years illustrates that an increasing number of journals published an above-average proportion of open access articles. 
In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a Creative Commons license in the same year. 
These journals were, on average, smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`). 
Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, and the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access).

When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements substantially contributed to the growth of open access over the years (Figure \ref{fig:results_overview}D). 
Examples of such journals include those with a scope on specific countries or regions, where also transformative agreements were implemented. 
For instance, in 2022, the Germany-based journals Zeitschrift für Erziehungswissenschaft and Zeitschrift für Politikwissenschaft, as well as the Scandinavian Political Studies adressing the Nordic countries, achieved an overall open access uptake of more than 90% just through transformative agreements. 
Despite the rise in transformative agreements, it is worth noting that other means of publishing open access in hybrid journals remained common. 
In total, `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_no_ta") |> nrow()` journals published open access articles from authors affiliated with institutions without transformative agreements in place, while `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_ta_oa") |> nrow()` journals published at least one open access article through a transformative agreement in the same year.

@@

```{r publisher_league_table}
publisher_league_tbl |>
  mutate(across(contains(c("perc")), ~ paste(round(., 3) * 100))) |>
  select(-contains("cum")) |>
  kbl(
    booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE, position = "H",
    caption ="Hybrid open access through transformative agreements market shares 2018-2022",
    col.names = c(
      "Publisher",
      "Total",
      "%",
      "Total",
      "%",
      "Total",
      "%",
      "Total",
      "%"
    )
  ) |>
  kable_styling(position = "center") |>
  add_header_above(c(
    " " = 1,
    "Hybrid journals" = 2,
    "Articles" = 2,
    "OA articles" = 2,
    "TA OA articles" = 2
  ))
```
